---
title: "Visual-maps of michigan"
author: "Adham Rishmawi"
format: html
editor: visual
---

## Libraries

```{r warning=FALSE,message=FALSE}
library(ggplot2)
library(tidyverse)
library(ggmap)
library(maps)
library(dplyr)
library(revgeo)
library(sp)
library(sf)
library(tmap)
```

## Data sets + Map Data

```{r, warning=FALSE,message=FALSE}
Population_to_unknown_lines <- read_csv("~/Mcgregor Project/Data material/CSV-extracts from sql/Population_to_unknown_lines.csv")
map_data <- map_data("state", region= "michigan")
```

## Config Data sets to compatibility state

```{r, warning=FALSE,message=FALSE}
#google's API key
#ggmap::register_google(key = "AIzaSyB_bVP1ljY7VcsQDSEnJXW4AHWwBqMKmZU")


#altering data to add long and lat for cities
## cities_lat <- cbind(geocode(as.character(Population_to_unknown_lines$name)),Population_to_unknown_lines)
# Assuming your dataset is stored in a variable called "mydata"
## write.csv(cities_lat, file = "C:\\Users\\nader\\OneDrive\\Documents\\Mcgregor Project\\Data material\\CSV-extracts from sql\\cities_lat.csv", row.names = TRUE)

cities_lat <- read.csv("~/Mcgregor Project/Data material/CSV-extracts from sql/cities_lat.csv")

```

## Plotting Michigan-failed

```{r}
# MI <- map_data("state", region="michigan")
# ggplot(MI, aes(x=long, y=lat))+geom_polygon()
```

```{r}
# filtered_cities <- cities_lat %>%
#   filter(lon >= -91, lon <= -80, lat >= 41, lat <= 49)
# 
# ggplot(filtered_cities, aes(x = lon, y = lat)) +
#   geom_polygon() +
#   coord_map() +
#   geom_point(data = filtered_cities, aes(x = lon, y = lat), color = "orange")

```

## Map using package Tmap

```{r, warning=FALSE,message=FALSE}
# Convert cities_lat to an sf spatial object
cities_sf <- st_as_sf(na.omit(cities_lat), coords = c("lon", "lat")) |>
  subset(select = -X)


#making a porportion column

# Create the proportion column with default value of 0.01
cities_sf$proportion <-  cities_sf$Total_unknown / cities_sf$Total_service


cities_sf$proportion <- as.numeric(cities_sf$proportion)
# Assuming you have a data frame named 'cities_sf' with the 'proportion' column

cities_sf$percentage <- cities_sf$proportion * 100



# Create a tmap object
map <- tm_shape(cities_sf) +
  tm_symbols(col = "percentage", size = 0.03, palette = "RdYlBu") + 
  tm_layout(bg.color = "white")

# Set the map's extent to focus on Michigan
map <- map +
  tm_view(bbox = c(-90, 41, -80, 49))

# Display the map
tmap_mode("view")
#("albatross")
map

```

## Alter map to display proportions above 0.00

```{r,  warning=FALSE,message=FALSE}

filtered_cities_sf <- cities_sf %>%
  filter(proportion != 0)

# Create a tmap object
map_nonzero <- tm_shape(filtered_cities_sf) +
  tm_layout(bg.color = "white") +
  tm_bubbles("population",col = "percentage",size = 0.03 )

# Display the map
tmap_mode("view")
#tmap_style("albatross")
map_nonzero +
tm_view(bbox = c(-90, 41, -80, 49))

```

## Make a map that displays unknown/total with heat map layout

```{r}
#FIGURE OUT HOW TO MAKE SPATIAL POINTS INTO POLYGONS

```

## alternative graph for top of page 4

```{r}
library(tmap)
library(sf)

# Convert cities_lat to an sf spatial object
cities_sf <- st_as_sf(na.omit(cities_lat), coords = c("lon", "lat")) |>
  subset(select = -X)

# Create the proportion column with default value of 0.01
cities_sf$proportion <- cities_sf$Total_unknown / cities_sf$Total_service
cities_sf$proportion <- as.numeric(cities_sf$proportion)

# Assuming you have a data frame named 'cities_sf' with the 'proportion' column
cities_sf$percentage <- cities_sf$proportion * 100

# Create a tmap object
map <- tm_shape(cities_sf) +
  tm_symbols(col = "population", size = "population", palette =  c("purple", "gold", "red","green"),breaks =  c(0, 1655, 4121, 10412, 700000))

# Set the map's extent to focus on Michigan
map <- map +
  tm_view(bbox = c(-90, 41, -80, 49))

# Customize the tmap theme to change the background color
map <- map +
  tm_layout(bg.color = "white")

# Display the map
tmap_mode("view")
map

tmap_save(map, "World_map.html")
```

The code snippet utilizes the tmap and sf libraries in R to create a map visualization of cities in Michigan. The map displays symbolized points representing different cities, where the size and color of the symbols correspond to the population data. The symbols are colored using a custom palette based on population ranges, with larger and differently colored symbols indicating areas with higher population densities. The map focuses on Michigan by setting the bounding box coordinates, and it has a white background. The resulting visualization provides an overview of the population distribution across cities in Michigan, highlighting urban areas with larger symbols colored in red and green.
