---
title: "pdsmi_expose"
author: "Adham Rishmawi"
format: html
editor: visual
---

## libraries + loading in data

```{r}
library(readxl)


data <- read_excel("C:\\Users\\nader\\OneDrive\\Documents\\Mcgregor Project\\EGLE docs\\pdsmi-summary.xlsx")
```

## making table

\# Count the number of '0' observations in each column

count_column1 \<- sum(data\$\`Unknown - No Information\` != 0)

count_column2 \<- table(data\$\`Known Lead\`)\["0"\]

count_column3 \<- table(data\$\`No Lead or GPCL\`)\["0"\]

\# Create a table with the results

result_table \<- data.frame(Column = c("Number of cities with 1 or more unknowns", "Number of cities with 1 or more lead lines", "Number of cities that have zero lead or galvanized lines"),

Count = c(count_column1, count_column2, count_column3))

\# Print the table

print(result_table)

```{r}

library(knitr)
library(kableExtra)
library(tinytex)

# Count the number of '0' observations in each column
count_column1 <- sum(data$`Unknown - No Information` != 0)
count_column2 <- table(data$`Known Lead`)["0"]
count_column3 <- table(data$`No Lead or GPCL`)["0"]

# Create a table with the results
result_table <- data.frame(Column = c("Number of cities with 1 or more unknowns", "Number of cities with 1 or more lead lines", "Number of cities that have zero lead or galvanized lines"),
                           Count = c(count_column1, count_column2, count_column3))

# Print the table
print(result_table)

# Save the table as a PDF
#tinytex::install_tinytex()  # Install TinyTeX if not already installed
kable(result_table, format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = "striped", full_width = FALSE) %>%
  save_kable(file = "result_table.pdf")





```

```{r}
data$human <- data$`Total Service Lines`|>
  as.numeric()
library(knitr)
library(kableExtra)

# Calculate the counts
unknown_per30 <- sum(data$`Unknown - No Information` / data$human > 0.30, na.rm = TRUE)
unknown_per40 <- sum(data$`Unknown - No Information` / data$human > 0.40, na.rm = TRUE)
unknown_per50 <- sum(data$`Unknown - No Information` / data$human > 0.50, na.rm = TRUE)

# Create a table
count_table <- data.frame(Unknown_Percentage = c("30%+", "40%+", "50%+"),
                          Count = c(unknown_per30, unknown_per40, unknown_per50))

# Format the table
count_table


```

```{r}
# Create a new column with the city names based on the unknown category
data$Unknown_Category <- ifelse(data$`Unknown - No Information` / data$human > 0.50, "Unknown > 50%",
                                ifelse(data$`Unknown - No Information` / data$human > 0.40, "Unknown > 40%",
                                       ifelse(data$`Unknown - No Information` / data$human > 0.30, "Unknown > 30%",
                                              "Unknown <= 30%")))


```

```{r}


# Filter the data for cities with the category "Unknown > 50%"
cities_50plus_unknown <- data$`fMuir Village of`[data$Unknown_Category == "Unknown > 50%"]|>
  na.omit()

# Filter the data for cities with the category "Unknown > 40%"
cities_40plus_unknown <- data$`fMuir Village of`[data$Unknown_Category == "Unknown > 40%"]|>
  na.omit()

# Filter the data for cities with the category "Unknown > 30%"
cities_30plus_unknown <- data$`fMuir Village of`[data$Unknown_Category == "Unknown > 30%"]|>
  na.omit()

# Filter the data for cities with the category "Unknown > 50%"
cities_30minus_unknown <- data$`fMuir Village of`[data$Unknown_Category == "Unknown <= 30%"]|>
  na.omit()




```

```{r}
# File path
file_path <- "C:\\Users\\nader\\OneDrive\\Documents\\Mcgregor Project\\Data material\\txt-csv-from-R\\data-50per.txt"

# Writing values to a text file
write.table(cities_50plus_unknown, file = file_path, sep = "\t", row.names = FALSE, col.names = FALSE)

# File path
file_path <- "C:\\Users\\nader\\OneDrive\\Documents\\Mcgregor Project\\Data material\\txt-csv-from-R\\data-40per.txt"

# Writing values to a text file
write.table(cities_40plus_unknown, file = file_path, sep = "\t", row.names = FALSE, col.names = FALSE)

# File path
file_path <- "C:\\Users\\nader\\OneDrive\\Documents\\Mcgregor Project\\Data material\\txt-csv-from-R\\data-30per.txt"

# Writing values to a text file
write.table(cities_30plus_unknown, file = file_path, sep = "\t", row.names = FALSE, col.names = FALSE)
# File path

file_path <- "C:\\Users\\nader\\OneDrive\\Documents\\Mcgregor Project\\Data material\\txt-csv-from-R\\data-30minusper.txt"

# Writing values to a text file
write.table(cities_30minus_unknown, file = file_path, sep = "\t", row.names = FALSE, col.names = FALSE)


```

## extracting data set to use in maps qmd

```{r}

write.csv(data, file = "C:\\Users\\nader\\OneDrive\\Documents\\Mcgregor Project\\Data material\\txt-csv-from-R\\pdsmi-unknown.csv", row.names = TRUE)
```
