---
title: "pdsmi_expose"
author: "Adham Rishmawi"
format: html
editor: visual
---

## libraries + loading in data

```{r}
library(readxl)

data <- read_excel("C:\\Users\\nader\\OneDrive\\Documents\\Mcgregor Project\\EGLE docs\\pdsmi-summary.xlsx")
```

## making table

\# Count the number of '0' observations in each column

count_column1 \<- sum(data\$\`Unknown - No Information\` != 0)

count_column2 \<- table(data\$\`Known Lead\`)\["0"\]

count_column3 \<- table(data\$\`No Lead or GPCL\`)\["0"\]

\# Create a table with the results

result_table \<- data.frame(Column = c("Number of cities with 1 or more unknowns", "Number of cities with 1 or more lead lines", "Number of cities that have zero lead or galvanized lines"),

Count = c(count_column1, count_column2, count_column3))

\# Print the table

print(result_table)

```{r}

library(knitr)
library(kableExtra)
library(tinytex)

# Count the number of '0' observations in each column
count_column1 <- sum(data$`Unknown - No Information` != 0)
count_column2 <- table(data$`Known Lead`)["0"]
count_column3 <- table(data$`No Lead or GPCL`)["0"]

# Create a table with the results
result_table <- data.frame(Column = c("Number of cities with 1 or more unknowns", "Number of cities with 1 or more lead lines", "Number of cities that have zero lead or galvanized lines"),
                           Count = c(count_column1, count_column2, count_column3))

# Print the table
print(result_table)

# Save the table as a PDF
#tinytex::install_tinytex()  # Install TinyTeX if not already installed
kable(result_table, format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = "striped", full_width = FALSE) %>%
  save_kable(file = "result_table.pdf")






```
